---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gardenr

<!-- badges: start -->
[![R-CMD-check](https://github.com/Chhr1s/gardenr/workflows/R-CMD-check/badge.svg)](https://github.com/Chhr1s/gardenr/actions)
<!-- badges: end -->

The goal of `{gardenr}` is to provide tools for general linear mixed effects model regression (GLMM) trees as implemented in the `{glmertree}` package.

## Installation

You can install the development version of `{gardenr}` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Chhr1s/gardenr")
```

## Example

Here is an example of using `cross_validate_it()` to help with hyperparameter tuning. 

```{r example, message=FALSE}
library(gardenr)
library(glmertree)
library(tidyverse)
```

## Simulate some data

Here's a function within the package to simulate some multilevel data. It has defaults, but these can be changed. use `?sim_multilevel()` for details.

```{r}
dat <- sim_multilevel()
```

## Split Data & Make CV Object

Wherever possible, I wanted to rely on good functions from the `{tidymodels}` framework. We can use `{rsample}` to spit the data and make a cross-validation object. 

```{r}
example_split <- rsample::initial_split(dat)
example_train <- rsample::training(example_split)
example_test  <-  rsample::testing(example_split)
cv <- rsample::vfold_cv(data = example_train, v = 10)
```

## Make a `Formula` object

This is the proposed GLMM tree formula. This is *not* a normal `formula` object, but a `Formula` object instead. See `?Formula::as.Formula` for an explanation of the differences.

```{r}
ex_formula <-
   Formula::as.Formula(
      'outcome ~ small_1 |
      (1 | id_vector) |
      small_2 + small_c_1 + small_c_2 + nuisance_1a + nuisance_c_1a'
      )
```

## Make a Tuning Grid

We can then use dials to make a tuning grid. Notice that the parameter objects for GLMM trees have already been made and are in `{gardenr}` (e.g., `alpha_par()`)

```{r}
tuning_grid <-
  dials::grid_max_entropy(
    maxdepth_par(maxdepth_min = 0L, maxdepth_max = 20L),
    alpha_par(alpha_min = 0.10, alpha_max = 0.001),
    trim_par(trim_min = 0.01, trim_max = 0.5),
    size = 10
  )

tuning_grid
```

## Fit the Model to the Cross-Validated Data

Here we fit the model to the cross-validated object.

```{r}
fitted <-
   cross_validate_it(
      cv_obj = cv,
      seed = 713,
      tuning_grid = tuning_grid,
      mod_formula = ex_formula, 
      cluster = id_vector
      )
```

## See Best Fitting Hyperparameters

```{r}
best_fit <- 
  fitted %>% 
  dplyr::arrange(rmse) 
best_fit
```

```{r}
best_fit_trained <- 
  lmertree(
    data = example_train, 
    formula = 
      ex_formula, 
    maxdepth = best_fit$maxdepth_par[1], 
    alpha = best_fit$alpha_par[1],
    trim = best_fit$trim_par[1], 
    cluster = id_vector,
    verbose = TRUE
  )
```

## See the Default Plot

```{r}
plot(best_fit_trained$tree)
```

## Get RMSE for unseen data

```{r}
example_test %>% 
  mutate(
    predictions =
      predict(
        best_fit_trained, 
        newdata = example_test, 
        allow.new.levels = TRUE
        )
    ) %>% 
  summarize(
    rmse_unseen = 
      rmse(observed_y = outcome, predicted_y = predictions)
    )
```

## Plot the Parameters in One Space

This can clarify what the difference is 

```{r}
summary(best_fit_trained$tree) %>%
  purrr::map_dfr(broom::tidy, .id = 'node') %>% 
  dplyr::mutate(node = factor(node)) %>% 
  ggplot(
    aes(
      x = term, 
      y = estimate, 
      color = node,
      ymin = estimate - 1.96*std.error, 
      ymax = estimate + 1.96*std.error
      )
    ) + 
  geom_pointrange(position = position_dodge(width = 0.5)) +
  theme_bw()
```

